---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# slurmworkflow

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/slurmworkflow)](https://CRAN.R-project.org/package=slurmworkflow)
[![R-CMD-check](https://github.com/EpiModel/slurmworkflow/workflows/R-CMD-check/badge.svg)](https://github.com/EpiModel/slurmworkflow/actions)
[![codecov](https://codecov.io/gh/EpiModel/slurmworkflow/branch/main/graph/badge.svg?token=eo2r0HeP8Z)](https://codecov.io/gh/EpiModel/slurmworkflow)
<!-- badges: end -->

**slurmworkflow** solves the issue of running multiple inter-dependant
jobs on a [slurm](https://slurm.schedmd.com/) equipped HPC without a long lived
job or a persistent SSH session.

A workflow is a predefined set of steps
([sbatch](https://slurm.schedmd.com/sbatch.html)s) to be executed on an HPC. By
default the steps are run sequentially. But **slurmworkflow** provides tools
for altering the execution order, allowing conditional execution of the steps
and loop like behavior.

## Installation

You can install the development version of **slurmworkflow** with:

``` r
remotes::install_github("EpiModel/slurmworkflow")
```

## Example

```{r example, eval = FALSE}
library(slurmworkflow)

```

